Read in the data
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggrepel)
nba <- read_csv('./nba.csv')
tail(nba)
```

What is the date range of this season?
```{r}
ggplot(data=nba, aes(x=gmDate)) +geom_point()
```

What is the range of points scored in a game by one team? Which team scored the most points in a game? Which team scored the fewest? 
```{r}
head(nba)
```

Calculate the regular season (`seasTyp`) record (win/loss) for each team (hint: use `group_by()` and `summarize()` or `count()`)
```{r}
scores <- nba %>% group_by(teamAbbr)  %>% count(teamRslt)
head(scores)

ggplot(scores, aes(y= n , x = teamAbbr, fill = teamRslt)) + geom_bar( stat='identity')
```

```{r}

```

Which team had the best record? Which had the worst record? 

If that was too easy, then calculate the record per month. Which team had the best month?

Calculate a new variable, which is win% for each team. Then plot win % vs. average points per game (`teamPTS`)
```{r}
nba%>% group_by(teamAbbr)%>% summarize(win_perc = sum(teamRslt == 'Win')/n(), avg_ppg = mean(teamPTS)) %>% ggplot(aes(x=avg_ppg , y= win_perc, label=teamAbbr)) + geom_point() +geom_label_repel()
```


If that's too easy, then plot win % vs. average point differential (`teamPTS - opptPTS`)
```{r}
nba%>% group_by(teamAbbr)%>% summarize(win_perc = sum(teamRslt == 'Win')/n(), dif_ppg = mean(teamPTS-opptPTS)) %>% ggplot(aes(x=dif_ppg , y= win_perc, label=teamAbbr)) + geom_point() +geom_label_repel()
```

Now calculate wins/losses each for home and away games (`teamLoc`). Who had the best home record? Who had the best away record? Can you plot Home vs. Away win % (or number of wins)? 
```{r}
nba %>% group_by(teamAbbr, teamLoc) %>% summarise(wins =  sum(teamRslt == 'Win')) %>%  ggplot() +geom_bar(aes(y= wins, x=teamAbbr, fill =teamLoc), stat= 'identity')
```
```{r}
nba %>% group_by(teamAbbr, teamLoc) %>% summarise(win =  sum(teamRslt == 'Win')) %>% spread(key=teamLoc, value= win) %>% summarise(normalized_proportion = 100*Home/(Home+Away)) %>% ggplot(aes(x=teamAbbr, y= normalized_proportion)) +geom_point()
```

Make that an interactive plotly plot. Try editing the `tooltip` text to show you the team for each data point.
```{r}
```

Now make a plot or two in plotly using the `plot_ly()` function. Try both univariate and multivariate plots. For help, see [this cheat sheet](https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf)
```{}

```

Hard challenge: use plotly or gganimate to show how the home and away win percentages changed from month to month.
```{r}
```

Can you run a PCA on the mean average win/loss data? Hint: to start you can use `summarize all` and `select_if`

Plot the first two PCs and see if you can color it by win/loss. What are the variables that load strongly onto PC1 and PC2?
```{r}
nba_mean=nba %>% filter(seasTyp=="Regular") %>% group_by(teamAbbr,teamRslt) %>% summarize_all(mean,na.rm=T) %>% select_if(~sum(is.na(.)) == 0)

```

more challenges: What about doing this on a game-by-game basis? What variables load strongly onto PC1 and PC2? Does it predict a win/loss for the focal team?

super hard challenge: Plot win percentages on a north american map using the team city as the location.
```{r}
```

only slightly super harder challenge: make your map interactive
```{r}

```